\sysfull

### 2018-04-12 paper/background.tex
added a todo and changed the title
editorial revisions to background

@@

Energy harvesting devices operate using energy extracted from sources such as radio frequency transmissions and solar energy. These devices elide tethered power or a battery, instead collecting energy into a capacitor, operating when sufficient energy accumulates, and upon depleting the energy, turning off and recharging.
%
%Batteryless operation has a number of important advantages, making intermittent computing an important research domain. Supplying power to billions~\cite{gartner_iot} of embedded computers using batteries is not sustainable. The European Commission estimates that more than 160 kilotons of consumer batteries enter the European Union annually~\cite{eu_batteries_2016}. Batteries are an environmental risk, are fragile, are limited in their number of charge/discharge cycles, and may require costly physical maintenance that is difficult or impossible deployed (e.g. in space~\cite{kicksat}). By contrast, super-capacitors are durable, promising millions of charge/discharge cycles~\cite[Sec. I]{ongaro_pwre_2012}. The main limitation in moving to capacitor-based energy storage is that capacitor energy density---and consequently operating discharge time---is orders of magnitude less than a battery. 
%
%Given current technology development, battery-less systems are best suited for very long-term sensing and monitoring where access to recharge is either prohibitive or impossible. These include battery-less image capture and processing~\cite{naderiparizi_rfid_2015}, animal monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015} and digestible~\cite{nadeau_naturebio_2017} sensors.
%
Many platforms enable intermittent, battery-less energy harvesting-based computation. For instance, computational RFIDs---open-source TI MSP430-based~\cite{wolverine} WISP~\cite{wisp5} (with its variants such as WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones such as~\cite{medusa_farsens_2017}. Other intermittently-powered platforms include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less phone~\cite{talla_imwut_2017}.

@@

\textcolor{red}{reference work by Intel}

@@

\begin{figure}
	\begin{subfigure}[t]{\linewidth}
		\centering \includegraphics[width=\columnwidth]{figures/crc_example}
		\caption{Simplified C code snippet of a CRC calculation from~\cite{hicks_mibench2_2016}: per-byte message division by a polynomial; \texttt{NV} denotes non-volatile variable declaration.}\label{fig:crc_example}
	\end{subfigure}
	\begin{subfigure}[t]{\linewidth}
		\centering \includegraphics[width=\columnwidth]{figures/crc_example_war}
		\caption{Execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 9) with stale reminder (line 3)}\label{fig:crc_example_war}
    \end{subfigure}
	\caption{\textbf{Code example demonstrating effect of write after read on volatile memory checkpointing.}}\label{fig:code_demo_incosistency}
\end{figure}

@@

\begin{table}
	\centering
	\footnotesize
	\begin{tabular}{|c|c|}
		\hline
		Model & Data Copied to/from NVRAM \\
		\hline\hline
		Mementos~\cite{mementos}                             & Reg. + Stack     \\
		DINO~\cite{dino}                                     & Reg. + WAR variables \\
		Chain~\cite{chain}                                   & PC   + Channel data\\
		Alpaca~\cite{alpaca}                                 & PC   + WAR variables \\
		Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Reg. (requires NV main memory) \\
		~\cite{intel} & XXX \\
		\hline
	\end{tabular}
	\caption{\textbf{Memory consistency enforcement overheads of intermittent execution models;} \emph{Reg.}: the entire register file, \emph{PC}: program counter, \emph{Channel data}: variables explicitly task-shared by the programmer, \emph{WAR variables}: variables involved in WAR dependences (\emph{NV}: non-volatile).}
	\label{table:chechpoint_comparison}
\end{table}

@@

\begin{figure}
\begin{subfigure}[t]{.49\columnwidth}
	\centering \includegraphics[width=\columnwidth]{figures/fram_write}
	\caption{Cost of \emph{write} operation}
\end{subfigure}%
\begin{subfigure}[t]{.49\columnwidth}
	\centering \includegraphics[width=\columnwidth]{figures/fram_read}
	\caption{Cost of \emph{read} operation}
\end{subfigure}
	\caption{\textbf{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write/read operation.}}\label{fig:framEnergy}
\end{figure}

>>>>

%\begin{figure}
\begin{wrapfigure}{t!}{0.5\textwidth}
	\begin{subfigure}[t]{\linewidth}
		\centering \includegraphics[width=\columnwidth]{figures/crc_example}
		\caption{Simplified C code snippet of a CRC calculation from~\cite{hicks_mibench2_2016}: per-byte message division by a polynomial; \texttt{NV} denotes non-volatile variable declaration.}\label{fig:crc_example}
	\end{subfigure}
	\begin{subfigure}[t]{\linewidth}
		\centering \includegraphics[width=\columnwidth]{figures/crc_example_war}
		\caption{Execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 9) with stale reminder (line 3)}\label{fig:crc_example_war}
	\end{subfigure}
	\caption{\textbf{Code example demonstrating effect of write after read on volatile memory checkpointing.}}\label{fig:code_demo_incosistency}
\end{wrapfigure}
%\end{figure}

@@

Energy harvesting devices operate using energy extracted from sources such as radio frequency transmissions and solar energy. These devices elide tethered power or a battery, instead collecting energy into a capacitor, operating when sufficient energy accumulates, and upon depleting the energy, turning off and recharging.
\TODO{We need a figure that shows what intermittent execution looks like.}
%
%Batteryless operation has a number of important advantages, making intermittent computing an important research domain. Supplying power to billions~\cite{gartner_iot} of embedded computers using batteries is not sustainable. The European Commission estimates that more than 160 kilotons of consumer batteries enter the European Union annually~\cite{eu_batteries_2016}. Batteries are an environmental risk, are fragile, are limited in their number of charge/discharge cycles, and may require costly physical maintenance that is difficult or impossible deployed (e.g. in space~\cite{kicksat}). By contrast, super-capacitors are durable, promising millions of charge/discharge cycles~\cite[Sec. I]{ongaro_pwre_2012}. The main limitation in moving to capacitor-based energy storage is that capacitor energy density---and consequently operating discharge time---is orders of magnitude less than a battery. 
%
%Given current technology development, battery-less systems are best suited for very long-term sensing and monitoring where access to recharge is either prohibitive or impossible. These include battery-less image capture and processing~\cite{naderiparizi_rfid_2015}, animal monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015} and digestible~\cite{nadeau_naturebio_2017} sensors.
%
Many platforms enable intermittent, battery-less energy harvesting-based computation. For instance, computational RFIDs---open-source TI MSP430-based~\cite{wolverine} WISP~\cite{wisp5} (with its variants such as WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones such as~\cite{medusa_farsens_2017}. Other intermittently-powered platforms include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less phone~\cite{talla_imwut_2017}.

@@

\textcolor{red}{reference work by Intel~\cite{baghsorkhi_cgo_2018}}

@@

@@

\begin{table}
	\centering
	\footnotesize
	\begin{tabular}{|c|c|}
		\hline
		Model & Data Copied to/from NVRAM \\
		\hline\hline
		Mementos~\cite{mementos}                             & Reg. + Stack     \\
		DINO~\cite{dino}                                     & Reg. + WAR variables \\
		Chain~\cite{chain}                                   & PC   + Channel data\\
		Alpaca~\cite{alpaca}                                 & PC   + WAR variables \\
		Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Reg. (requires NV main memory) \\
		Region Formation~\cite{baghsorkhi_cgo_2018} & \textcolor{red}{XXX} \\
		\hline
	\end{tabular}
	\caption{\textbf{Memory consistency enforcement overheads of intermittent execution models;} \emph{Reg.}: the entire register file, \emph{PC}: program counter, \emph{Channel data}: variables explicitly task-shared by the programmer, \emph{WAR variables}: variables involved in WAR dependences (\emph{NV}: non-volatile).}
	\label{table:chechpoint_comparison}
\end{table}

@@

\begin{figure}
\begin{subfigure}[t]{.49\columnwidth}
	\centering \includegraphics[width=\columnwidth]{figures/fram_write}
	\caption{Cost of \emph{write} operation}
\end{subfigure}%
\begin{subfigure}[t]{.49\columnwidth}
	\centering \includegraphics[width=\columnwidth]{figures/fram_read}
	\caption{Cost of \emph{read} operation}
\end{subfigure}
	\caption{\textbf{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write/read operation.} \textcolor{red}{figures need to be merged into one}}\label{fig:framEnergy}
\end{figure}
